---
title: "Applied Web Scraping"
author: "Tobias Rüttenauer"
date: "March 24, 2023"
output_dir: docs
output: 
  html_document:
    theme: flatly
    highlight: haddock
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
    toc_depth: 2
theme: united
bibliography: sicss-spatial.bib
link-citations: yes
---

### Required packages

```{r, message = FALSE, warning = FALSE, results = 'hide'}
pkgs <- c("sf", "gstat", "mapview", "dplyr",
          "rvest", "xml2",
          "nomisr", "osmdata", "tidyr") 
lapply(pkgs, require, character.only = TRUE)

```

### Session info

```{r}
sessionInfo()

```

```{r}
# Create subdir for data if not exists
dn <- "_data"
ifelse(dir.exists(dn), "Exists", dir.create(dn))
```


# Web Scraping

This example is inspired by @Biegert.2023.

## Football Stats

Let's try to get some football player stats from [FBREF](https://fbref.com/en/). On the home page, we see a list of trending player pages: [https://fbref.com/en/]() Let's see if we can get this information into R.

![[https://fbref.com/en/](https://fbref.com/en/)](fig/fbref2.png)

Inspecting the head shots of players give is the following xpath: `//*[@id="players"]/div[3]`

```{r eval=FALSE}
# Scrape the website
url <- "https://fbref.com/en/"
parsed <- read_html(url)

# Save xml (so we have the website locally)
write_xml(parsed, file = "_data/fbplayers.xml")
```

```{r}
# Read back in
parsed <- read_html("_data/fbplayers.xml")

# Select the desired element
parsed.sub <- html_elements(parsed, xpath =  '//*[@id="players"]/div[3]')
trending.players <- html_elements(parsed.sub, "a")

trending.players
```

We have extracted the set of players with the respective links. Now, we separate these into the names with `html_text2()` and the links using `html_attrs()`.

```{r}
# Data frame with names and links of players
trending.players.df <- data.frame(names = html_text2(trending.players),
                       links = unlist(html_attrs(trending.players)))

# Extend the links
trending.players.df$links <- paste0("https://fbref.com/", trending.players.df$links)

head(trending.players.df)

```

## Exercise

- Can you get some player characteristics for Lionel Messi?

- Can you wrap this into a loop for multiple players? __USE `sys.sleep()` TO PAUSE!__

```{r eval=FALSE, echo = FALSE}

### We will loop over these popular players

# I set this to the first three only. Obviously, you could do more

ids <- 1:3

for(i in ids){
  
  # Get individual ids
  player.name <- trending.players.df$names[i]
  player.link <- trending.players.df$links[i]
  
  # parse the website
  parsed.tmp <- read_html(player.link, as.data.frame = TRUE, stringAsFactors = TRUE )
  
  # Extract tables from the website
  tables.tmp <- html_nodes(parsed.tmp, "table")
  
  # Look at attributes of these tables
  attr.df <- data.frame(do.call(rbind, html_attrs(tables.tmp)))
  
  # Select the table "stats_standard_dom_lg"
  oo <- which(attr.df$id == "stats_standard_dom_lg")
  table.tmp <- html_table(tables.tmp[[oo]], fill = TRUE)
  
  # Clean the table (first row is still part of the names)
  names <- janitor::make_clean_names(paste0(names(table.tmp), table.tmp[1,]))
  names(table.tmp) <- names
  table.tmp <- table.tmp[-1,]
  
  # Drop summary lines (they all contain non-numeric values for age)
  oo <- which(!is.na(as.numeric(table.tmp$age)))
  table.tmp <- table.tmp[oo,]
  
  # Add player name
  table.tmp <- data.frame(player = player.name, table.tmp)
  
  # Combine into a common results frame
  if(i == 1){
    player_stats.df <- table.tmp
  }else{
    player_stats.df <- plyr::rbind.fill(player_stats.df, table.tmp) 
  }
  
  ### Slow the loop down!!!
  Sys.sleep(3)
}

```

Now we have a time-series object of the most trending player's seasonal performances.

```{r eval = FALSE}
head(player_stats.df)
```


# API

This is from a short course [Introduction to geospatial data and analysis](https://ruettenauer.github.io/SICSS-Spatial2022/index.html)

## London shapefile (polygon)

Lets get some administrative boundaries for London from the [London Datastore](https://data.london.gov.uk/dataset/statistical-gis-boundary-files-london). 
```{r}
# Download zip file and unzip
tmpf <- tempfile()
boundary.link <- "https://data.london.gov.uk/download/statistical-gis-boundary-files-london/9ba8c833-6370-4b11-abdc-314aa020d5e0/statistical-gis-boundaries-london.zip"
download.file(boundary.link, tmpf)
unzip(zipfile = tmpf, exdir = paste0(dn))
unlink(tmpf)

# This is a shapefile
# We only need the MSOA layer for now
msoa.spdf <- st_read(dsn = paste0(dn, "/statistical-gis-boundaries-london/ESRI"),
                     layer = "MSOA_2011_London_gen_MHW" # Note: no file ending
                     )
head(msoa.spdf)

```

This looks like a conventional `data.frame` but has the additional column `geometry` containing the coordinates of each observation. `st_geometry()` returns only the geographic object and `st_drop_geometry()` only the `data.frame` without the coordinates. We can plot the object using `mapview()`.

```{r}
mapview(msoa.spdf[, "POPDEN"])

```

## Census API (admin units)

Now that we have some boundaries and shapes of spatial units in London, we can start looking for different data sources to populate the geometries.

A good source for demographic data is for instance the 2011 census. Below we use the nomis API to retrieve population data for London, See the [Vignette](https://cran.r-project.org/web/packages/nomisr/vignettes/introduction.html) for more information (Guest users are limited to 25,000 rows per query). Below is a wrapper to avoid some mess with sex and urban-rural cross-tabs.

```{r}
### For larger request, register and set key
# Sys.setenv(NOMIS_API_KEY = "XXX")
# nomis_api_key(check_env = TRUE)



# Get London ids
london_ids <- msoa.spdf$MSOA11CD

### Get key statistics ids
# select required tables (https://www.nomisweb.co.uk/sources/census_2011_ks)
# Let's get KS201EW (ethnic group) and KS402EW (housing tenure)

```

From the Nomis website [https://www.nomisweb.co.uk/sources/census_2011_ks](https://www.nomisweb.co.uk/sources/census_2011_ks), we can get the external identifiers that we want to pull: KS201EW (ethnic group) and KS402EW (housing tenure). Unfortunately, internal keys differ (for what reason ever...). 

```{r}
# Get table of IDs and description
x <- nomis_data_info()
head(x)
```

So we have to find the internal ids of KS201EW and KS402EW.

```{r}
# Get internal ids
stats <- c("KS201EW", "KS402EW")
oo <- which(grepl(paste(stats, collapse = "|"), x$name.value))
ksids <- x$id[oo]
ksids # This are the internal ids
```

Finally, we can start pulling the information from the Nomis census APi! Below is a wrapper to avoid some mess with sex and urban-rural cross-tabs. Some characteristics can be pulled separately for urban-rural or by sex - In that case we have to specify it, but we cannot specify the option if a variable does not have urban-rural or sex differentials. The wrapper below also creates a codebook with variable descriptions.

```{r}
### look at meta information
q <- nomis_overview(ksids[1])
head(q)
a <- nomis_get_metadata(id = ksids[1], concept = "GEOGRAPHY", type = "type")
a # TYPE297 is MSOA level

b <- nomis_get_metadata(id = ksids[1], concept = "MEASURES", type = "TYPE297")
b # 20100 is the measure of absolute numbers


### Query data in loop over the required statistics
for(i in ksids){

  # Determin if data is divided by sex or urban-rural
  nd <- nomis_get_metadata(id = i)
  if("RURAL_URBAN" %in% nd$conceptref){
    UR <- TRUE
  }else{
    UR <- FALSE
  }
  if("C_SEX" %in% nd$conceptref){
    SEX <- TRUE
  }else{
    SEX <- FALSE
  }

  # make data request
  if(UR == TRUE){
    if(SEX == TRUE){
      tmp_en <- nomis_get_data(id = i, time = "2011", 
                               geography = london_ids, # replace with "TYPE297" for all MSOAs
                               measures = 20100, RURAL_URBAN = 0, C_SEX = 0)
    }else{
      tmp_en <- nomis_get_data(id = i, time = "2011", 
                               geography = london_ids, # replace with "TYPE297" for all MSOAs
                               measures = 20100, RURAL_URBAN = 0)
    }
  }else{
    if(SEX == TRUE){
      tmp_en <- nomis_get_data(id = i, time = "2011", 
                               geography = london_ids, # replace with "TYPE297" for all MSOAs
                               measures = 20100, C_SEX = 0)
    }else{
      tmp_en <- nomis_get_data(id = i, time = "2011", 
                               geography = london_ids, # replace with "TYPE297" for all MSOAs
                               measures = 20100)
    }

  }

  # Append (in case of different regions)
  ks_tmp <- tmp_en

  # Make lower case names
  names(ks_tmp) <- tolower(names(ks_tmp))
  names(ks_tmp)[names(ks_tmp) == "geography_code"] <- "msoa11"
  names(ks_tmp)[names(ks_tmp) == "geography_name"] <- "name"

  # replace weird cell codes
  onlynum <- which(grepl("^[[:digit:]]+$", ks_tmp$cell_code))
  if(length(onlynum) != 0){
    code <- substr(ks_tmp$cell_code[-onlynum][1], 1, 7)
    if(is.na(code)){
      code <- i
    }
    ks_tmp$cell_code[onlynum] <- paste0(code, "_", ks_tmp$cell_code[onlynum])
  }

  # save codebook
  ks_cb <- unique(ks_tmp[, c("date", "cell_type", "cell", "cell_code", "cell_name")])

  ### Reshape
  ks_res <- tidyr::pivot_wider(ks_tmp, id_cols = c("msoa11", "name"),
                               names_from = "cell_code",
                               values_from = "obs_value")

  ### Merge
  if(i == ksids[1]){
    census_keystat.df <- ks_res
    census_keystat_cb.df <- ks_cb
  }else{
    census_keystat.df <- merge(census_keystat.df, ks_res, by = c("msoa11", "name"), all = TRUE)
    census_keystat_cb.df <- rbind(census_keystat_cb.df, ks_cb)
  }

}


# Descirption are saved in the codebook
head(census_keystat_cb.df)

### Merge with MSOA geometries above
msoa.spdf <- merge(msoa.spdf, census_keystat.df, 
                   by.x = "MSOA11CD", by.y = "msoa11", all.x = TRUE)

```

Now we can, for instance, plot the spatial distribution of ethnic groups.

```{r}

msoa.spdf$per_white <- msoa.spdf$KS201EW_100 / msoa.spdf$KS201EW0001 * 100

mapview(msoa.spdf[, "per_white"])

```





# References
